Exploratory Data Analysis of white wine data by Dheeraj R Kondati
========================================================



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(GGally)


```

```{r echo=FALSE, Load_the_Data}
setwd("C:/Users/Dheeraj/Desktop/nano/p4/project")
white<-read.csv("wineQualityWhites.csv")

```

#I.DATASET INFORMATION
##1.CITATION:
This dataset is public available for research. The details are described in [Cortez et al., 2009]. 
  
P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
Modeling wine preferences by data mining from physicochemical properties.
In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016
[Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf
[bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib

##2.TITLE:
Wine Quality-White Wine

##3.ABOUT:
This data set contains 4,898 white wines with 11 variables on quantifying the chemical properties of each wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

##4.Number of Instances: 4898

##5.Attributes:
There are 11 input attributes and one output attribute and one serial number attribute.

###Input Attributes
Based on physicochemical tests
1-fixed acidity (tartaric acid - g / dm^3)
2-volatile acidity (acetic acid - g / dm^3)
3-citric acid (g / dm^3)
4-residual sugar (g / dm^3)
5-chlorides (sodium chloride - g / dm^3
6-free sulfur dioxide (mg / dm^3)
7-total sulfur dioxide (mg / dm^3)
8-density (g / cm^3)
9-pH
10-sulphates (potassium sulphate - g / dm3)
11-alcohol (% by volume)

###Output variable (based on sensory data) 
12 - quality (score between 0 and 10)

###Serial Number Attribute
X-1-4898

##6.Attribute Information
1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)<br/>
2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste<br/>
3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines<br/>
4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet<br/>
5 - chlorides: the amount of salt in the wine<br/>
6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine<br/>
7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine<br/>
8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content<br/>
9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale<br/>
10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant<br/>
11 - alcohol: the percent alcohol content of the wine<br/>

Output variable (based on sensory data): <br/>
12 - quality (score between 0 and 10)<br/>

#II.OVERVIEW OF THE DATA

Dimensions of Data

```{r echo=FALSE,message=FALSE, warning=FALSE ,Univariate_Plots}
dim(white)

```

The first 5 rows of the dataset

```{r echo=FALSE}
head(white)

```

Summary of datatypes of the attributes

```{r echo=FALSE}
str(white)

```

From the summary above,we can observe that quality and X is an integer class,while others are stored as the numeric class<br/>
<br/>

Overall Summary of all attributes

```{r echo=FALSE}
summary(white)

```

From the summary, we can observe that there are a few attributes with outliers,namely residual.sugar and free.sulphur.dioxide.

#III Single Variable Analysis
##Overview of each variable

```{r echo=FALSE, message=FALSE, warning=FALSE}
p1<-qplot(white$fixed.acidity)

p2<-qplot(white$volatile.acidity)##oultiers

p3<-qplot(white$citric.acid)##outliers

p4<-qplot(white$residual.sugar)##outliers

p5<-qplot(white$chlorides)##outliers

p6<-qplot(white$free.sulfur.dioxide)##long tail

p7<-qplot(white$total.sulfur.dioxide)##outliers

p8<-qplot(white$density)#outliers

p9<-qplot(white$pH)#long tail

p10<-qplot(white$sulphates)##outliers

p11<-qplot(white$alcohol)#outliers
grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,ncol=4)





```

Outliers and long tails can be seen from the intital overview,a few of the outliers are cleaned

```{r echo=FALSE, message=FALSE, warning=FALSE}
white<-subset(white,white$fixed.acidity > quantile(white$fixed.acidity,probs=0.01)& white$fixed.acidity < quantile(white$fixed.acidity,probs = 0.99))
white<-subset(white,white$volatile.acidity > quantile(white$volatile.acidity,probs=0.01)& white$volatile.acidity < quantile(white$volatile.acidity,probs = 0.99))
white<-subset(white,white$chlorides > quantile(white$chlorides,probs=0.01)& white$chlorides < quantile(white$chlorides,probs = 0.99))
white<-subset(white,white$total.sulfur.dioxide > quantile(white$total.sulfur.dioxide,probs=0.01)& white$total.sulfur.dioxide < quantile(white$total.sulfur.dioxide,probs = 0.99))
white<-subset(white,white$density > quantile(white$density,probs=0.01)& white$density < quantile(white$density,probs = 0.99))


```

```{r echo=FALSE, message=FALSE, warning=FALSE}
p1<-qplot(white$fixed.acidity)

p2<-qplot(white$volatile.acidity)##oultiers

p3<-qplot(white$citric.acid)##outliers

p4<-qplot(white$residual.sugar)##outliers

p5<-qplot(white$chlorides)##outliers

p6<-qplot(white$free.sulfur.dioxide)##long tail

p7<-qplot(white$total.sulfur.dioxide)##outliers

p8<-qplot(white$density)#outliers

p9<-qplot(white$pH)#long tail

p10<-qplot(white$sulphates)##outliers

p11<-qplot(white$alcohol)#outliers
grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,ncol=4)

```

##Closer look at the data

###a)Acidity

```{r echo=FALSE,warning=FALSE,message=FALSE}
p1<-qplot(x=fixed.acidity,data=white,binwidth=0.5,fill=I("sienna4"),col=I("black"))+scale_x_continuous(limits=c(4,10),breaks=seq(4,10,2))
p2<-qplot(x=volatile.acidity,data=white,binwidth=0.05,fill=I("sienna4"),col=I("black"))+scale_x_continuous(limits=c(0.05,0.65),breaks=seq(0.05,0.65,0.1))

grid.arrange(p1,p2,ncol=2)



```

From the plots, we can observe a slight skewness in both the acidity values, most of the values for fixed acidity is between 6-8 and for volatile acidity is between 0.2 to 0.3. There are outliers which are eliminated from the plots.

###b)Sulphurdioxide(SO2) content

```{r echo=FALSE,warning=FALSE,message=FALSE}
p3<-qplot(x=free.sulfur.dioxide,data=white,binwidth=5,fill=I("sienna4"),col=I("black"))+scale_x_continuous(limits=c(2,75))
p4<-qplot(x=total.sulfur.dioxide,data=white,binwidth=30,fill=I("sienna4"),col=I("black"))+scale_x_continuous(limits=c(2,300))
grid.arrange(p3,p4,ncol=2)

```

Both the free sulphur dioxide and total sulphur dioxide have long tails,these data points have been eiminated from the graph.Most of the data points for total SO2 are between 100-160 and for free SO2, they are between 20-40.

###Chemical Compositions

```{r echo=FALSE,warning=FALSE,message=FALSE}
p5<-qplot(x=citric.acid,data=white,binwidth=0.01,fill=I("sienna4"),col=I("black"))+scale_x_continuous(limits=c(0,0.7),breaks=seq(0,0.7,0.05))

p6<-qplot(x=residual.sugar,data=white,binwidth=1,fill=I("sienna4"),col=I("black"))+scale_x_continuous(limits=c(0,20),breaks=seq(0,20,5))

p7<-qplot(x=white$chlorides,xlim = c(0.020,0.1),fill=I("sienna4"),col=I("black"))

p8<-qplot(x=sulphates,data=white,binwidth=0.1,fill=I("sienna4"),col=I("black"))+scale_x_continuous(limits=c(0.2,1.1),breaks=seq(0.2,1.1,0.1))

grid.arrange(p5,p6,p7,p8,ncol=2)


```

Residual.Sugar is heavily skewed with nearly half of the data within 0-5, White chlorides have most of the data between 0.025-0.07,Sulphates have a long tail and citric.acid shows an unusual peak near 0.5.

###pH

```{r echo=FALSE,warning=FALSE,message=FALSE}
qplot(x=pH,data=white,fill=I("sienna4"),col=I("black"))


```

Fairly Normal Distribution of pH values,most of the values between 3.1 and 3.3

###Density

```{r echo=FALSE,warning=FALSE,message=FALSE}
qplot(x=density,data=white,binwidth=0.001,fill=I("sienna4"),col=I("black"))+scale_x_continuous(limits=c(0.987,1.002),breaks=seq(0.987,1.002,0.01))


```

Density data has outlier data points and a long tail,the plot does not show the tail.

###Quality

```{r echo=FALSE,warning=FALSE,message=FALSE}
qplot(white$quality,binwidth=1,fill=I("sienna4"),col=I("black"))+scale_x_continuous(breaks=seq(3,9,1))

```

Most of the wines are rated at 5 and 6

### What is the structure of your dataset?
There are 4898 instances of white wines with 11 independent attributes for each wine,each wine is rated on a scale of 1-10 which is the dependent variable as it is influenced by the attributes.

### What is/are the main feature(s) of interest in your dataset?
Main feature of interest of the dataset is to observe how each attribute influences the quality of the alcohol.But from knowledge we can say that alcohol could be a driving feature as it is the most important chemical component of wine.A model could be built based on the relation between the attributes and quality which could predict the quality given attributes of a new wine.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Based on the initial view of the data, any of the variables could influence the quality so no variable can be ruled out yet.Certainly Alcohol will be an important feature.On observation of correlations between attributes and quality a deduction could be made in the attribute space.

### Did you create any new variables from existing variables in the dataset?
No new variables have been created.But new variables could be created which would segment the variables to create better visualisations.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

There are a few outliers in the dataset which have been eliminated.Mainly the datapoints above the 99 percentile and below the 0.01 percentile have been removed from the following attributes fixed.acidity, volatile.acidity,chlorides,total.sulfur.dioxide,density.


# Bivariate Plots Section

###Correlations Plot


```{r echo=FALSE, Bivariate_Plots}
ggpairs(white[sample.int(nrow(white),1000),])


```

Strong correlations observed.<br/>
1.Alcohol vs density -0.813<br/>
2.density vs residual sugar 0.835<br/>
3.density vs Total sulfur dioxide 0.535<br/>
4.Residual sugar vs Total sulfur dioxide 0.414<br/>

We are more interested in checking the influence of attributes on quality,lets look at those plots first.

### Alcohol Influence on quality

Wine has relaxing effects,from our knowledge we know that Alcohol is the reason.Lets check how alcohol levels influence quality of the wine.

```{r echo=FALSE, message=FALSE, warning=FALSE}
whitef<-white
whitef$quality<-as.factor(white$quality)
biplot<-function(p){
  plt<-ggplot(aes(x=quality , y = p),data=white)+
    geom_jitter(alpha=1/2,shape=1)+
    geom_smooth(method='lm',color='red')+
    scale_x_continuous(breaks=seq(3,9,1))
  return(plt)
  
}
biplot(white$alcohol)

```

It can be seen that as the alcohol level increases,the wine rating increases

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=quality,y=alcohol),data=white)+
  geom_jitter(alpha=1/2,shape=21)+
  
  geom_line(stat='summary',fun.y=mean,color='blue')+
  geom_line(stat='summary',fun.y=median,color='red')+
  
  scale_x_continuous(breaks=seq(3,9,1))

```

The lines in the plot are the median and mean alcohol levels across each level of quality.It can be seen that there is an increasing trend.

```{r echo=FALSE, message=FALSE, warning=FALSE}
boxplt<-function(q){
  plt2<-ggplot(aes(x=quality,y=q),data=whitef)+
    geom_boxplot()+
    geom_hline(yintercept=median(q),alpha=1,linetype=2,size=1,color='red')
  return(plt2)
}
boxplt(white$alcohol)



```

The red line is the overall median of alcohol values,higher alcohol content clearly signifies better wine.

###citirc acidinfluence on quality

```{r echo=FALSE, message=FALSE, warning=FALSE}
biplot(white$citric.acid)
boxplt(white$citric.acid)

```

From the plots, it can be seen that there is no correlation between citric acid levels and quality and most of the wines have pH values at 0.2-0.3

###sulphates influence on quality

```{r echo=FALSE, message=FALSE, warning=FALSE}
biplot(white$sulphates)
boxplt(white$sulphates)

```

###acidity on quality

```{r echo=FALSE, message=FALSE, warning=FALSE}
biplot(white$fixed.acidity)
boxplt(white$fixed.acidity)


```

```{r echo=FALSE, message=FALSE, warning=FALSE}
biplot(white$volatile.acidity)
boxplt(white$volatile.acidity)

```

###chlorides

```{r echo=FALSE, message=FALSE, warning=FALSE}
biplot(white$chlorides)
boxplt(white$chlorides)

```

###residual.sugar

```{r echo=FALSE, message=FALSE, warning=FALSE}
biplot(white$residual.sugar)
boxplt(white$residual.sugar)

```

###pH

```{r echo=FALSE, message=FALSE, warning=FALSE}
biplot(white$pH)
boxplt(white$pH)

```

###Summary of attributes on Quality
1.Citric acid,sulphates and fixed acidity have negligible influence on Quality.<br/>
2.As volatile acidity,chlorides and resudial sugar decrease the quality increases.<br/>
3.As pH value increases, Quality increases.

##Looking at other strong correlations

###Alcohol vs Density

From the scatterplot of correlations
ALcohol and Density have high correlation of -0.8.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=alcohol,y=density),data = white)+
  geom_jitter(alpha=1/2)+
  geom_smooth(method='lm',color='red')

```

The plot shows that there is a decreasing trend, as alcohol increases density decreases.

###density vs residual sugar

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=residual.sugar,y=density),data = white)+
  geom_jitter(alpha=1/2)+
  geom_smooth(method='lm',color='red')


```

Clearly as sugar level increases the wine becomes denser.


###density vs Total Sulfur dioxide

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=total.sulfur.dioxide,y=density),data = white)+
  geom_jitter(alpha=1/2)+
  geom_smooth(method='lm',color='red')


```

Denser wines have more total sulfur dioxide.




### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
Alcohol,density,residual sugar and total sufur dioxide have been found to be either strongly correlated to one another or have influenced quality.<br/>
Summary</br>
1.Citric acid,sulphates and fixed acidity have negligible influence on Quality.<br/>
2.As volatile acidity,chlorides and resudial sugar decrease the quality increases.<br/>
3.As pH value increases, Quality increases.<br/>


### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

There are a few strongly correlated attributes.<br/>
Summary<br/>
1.As alcohol increases density decreases.The smoothing line slopes downwards from left to right which supports this conslusion.Also the median values of higher quality wines are much more than the overall median.<br/>
2.As sugar level increases the wine becomes denser.<br/>
3.Denser wines have more total sulfur dioxide.<br/>



### What was the strongest relationship you found?
density and sugar are very highly correlated,Alcohol and density are also correlated but negatively. The plots and correlation values support this conclusion as shown above.

# Multivariate Plots Section



###Alcohol vs density quality

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x=alcohol,y=density, color=quality),data = subset(whitef,whitef$density<1.005))+
  geom_jitter(alpha=1)+
  scale_color_manual(values=c('red', 'blue', 'green', 'orange', 'pink', 'purple', 'black'))+
  facet_wrap(~quality)

```

From the plot,it is clarly seen that as alcohol increases and density decreases ,the wine gets a better rating


###Density vs residual sugar and alcohol

```{r echo=FALSE, message=FALSE, warning=FALSE}
white$alc_buckets<-cut(white$alcohol,c(8,10,12,14.10))
whitef$alc_buckets<-cut(whitef$alcohol,c(8,10,12,14.10))
ggplot(aes(x=residual.sugar,y=density, color=alc_buckets),data = subset(whitef,whitef$density<1.005))+
  geom_jitter(alpha=1)+
  scale_color_manual(values=c('red', 'blue', 'green'))
  


```

Highly dense wines have less alcohol and more sweetness.<br/>
Wines with higher alcohol contents have less density and sweetness.

###Density vs Total sulfur dioxide vs alcohol

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=total.sulfur.dioxide,y=density, color=alc_buckets),data = subset(whitef,whitef$density<1.005))+
  geom_jitter(alpha=1/4)+
  scale_color_manual(values=c('red', 'blue', 'green'))


```

3 distinct patches have been observed.Less denser wines have less alcohol content and low so2 levels.

#Fitting a linear model

```{r echo=FALSE, message=FALSE, warning=FALSE}
b<- lm(quality~.,data=white)
summary(b)

```



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
The relationships which I observed were 1)Alcohol,density and quality 2)Density,Sugar,Alcohol 3)density,total sulfur dioxide and Alcohol<br/>
1)Alcohol,density and quality : As alcohol increases density decreases,the wine following this trend gets a better rating,i.e a poorly rated wine has low alcohol content and high density<br/>
2)Density,Sugar,Alcohol : Highly dense wines have less alcohol and more sweetness.Wines with higher alcohol contents have less density and sweetness.
3)density,total sulfur dioxide and Alcohol: Less denser wines have less alcohol content and low so2 levels.

### Were there any interesting or surprising interactions between features?

A surprising interaction was that though the correlation between alcohol and quality is less it has significant effect on the rating as seen from exploratory data analysis.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.
A linear model was fit to the data.The coefficients of the variables clearly show the importance of variables which we have been looking at.There are few attributes getting surprisingly heavier weights.The model is very basic and has high error.


------

# Final Plots and Summary



### Plot One

```{r echo=FALSE, Plot_One}
ggplot(aes(x=quality,y=alcohol),data=whitef)+
  xlab('quality of wine(0-10)')+
  ylab('alcohol (% by volume)') +
  ggtitle("Comparing Alcohol contents between all qualities of the wines")+  
  geom_boxplot()+
  geom_hline(yintercept=median(whitef$alcohol),alpha=1,linetype=2,size=1,color='red')


```

### Description One
This plot clearly shows the influence of alcohol on the wine quality.The correlation between alcohol and quality though not really high surely has great significance in determining quality of wine.I have first converted the data type of quality from continuous to discrete ,this helps in visualising the boxplot.The geom_boxplot function was used to plot boxplots across the different qualities.I then used the geom hline function to plot the overall median line for the alcohol data<br/>
An interesting finding for me was, that the alcohol level shows a decreasing trend as quality increases but the trend stops and reverses,as alcohol level increases beyond a certain point,the quality also increases.The decreasing trend could be because of two reasons, that the alcohol level reached a minima and showed decremental effects on wine quality so alcohol contents had to be increased,  or the quality scale is so wide that attributes between two steps of quality(in this case 4 and 5) are hard to discern and were rated differently.<br/>
If we continue our reasoning with the first option,it could be that once the alcohol levels reaches a maxima it would starts to show negative effects on quality subsequently alcohol contents would be reduced to find an optima.


### Plot Two

```{r echo=FALSE, Plot_Two}
ggplot(aes(x=alcohol,y=density, color=quality),data = subset(whitef,whitef$density<1.005))+
  geom_jitter(alpha=1)+
  ggtitle("Scatterplot of density vs alcohol with color set by Quality")+
  xlab("alcohol (% by volume)")+
  ylab("density (g/cm^3)")+
  scale_color_manual(values=c('red', 'blue', 'green', 'orange', 'pink', 'purple', 'black'))+
  facet_wrap(~quality)


```


### Description Two
This plot is a good visualisation to see the influence of alcohol and density on the quality of wine.To develop this visualisation,I changed the datatype of quality from continuous to discrete and used different colors for the qualities.Also the plot is faceted based on quality.<br>
I can see from the plots that there is a cluster of points concentrated in different parts of the graph for each quality.The cluster eventually moves from top left to bottom right.From this I can infer that as alcohol content increased and density became lesser,the quality of the wine increased.</br>
The cluster is less denser for the higer qualities as there are only a few data points which are rated high.<br/>
The plot shows a trend which cannot be clearly seen from correlations due to less datapoints of wines with higher ratings.Exploratory Data Analysis helped uncover this trend.


### Plot Three

```{r echo=FALSE, Plot_Three}
ggplot(aes(x=residual.sugar,y=density, color=alc_buckets),data = subset(whitef,whitef$density<1.005))+
  geom_jitter(alpha=1)+
  ggtitle("Scatterplot of density vs residual sugar with color set by alcohol level")+
  xlab("residual sugar  (g / dm^3)")+
  ylab("density (g/cm^3)")+
  scale_color_manual(values=c('red', 'blue', 'green'))


```

### Description Three
This plot is a clear visualisation of the correlation between density,residual sugar and alcohol level.To obtain the coloring,I segmented the alcohol levels into 3 buckets to visualise low medium and high levels of alcohol contents.From the plot there are clearly different clusters as we can observe from the coloring.The green patch at the bottom left are all the wines with low density and residual sugars,all wines with high alcohol content fall in this region.The blue region consists of wines with medim alcohol levels,the density and sugar levels are also moderate.The red patch consists of all the wines with low alcohol content and high density and residual sugar values.<br/>
From my study of wine preparation,the residual sugar is the sugar left in the wine after fermentation.This plot is interesting as it tells us that wines with low alcohol are sweeter than wines with high alcohol levels.Also that less denser wines are going to taste bitter.<br/>


------

# Reflection
The process of Exploratory Data Analysis of the wine data has introduced me to various tools in R which helped me in visualising data quickly and clearly.EDA helped me in gaining insight into trends and patterns present in the data.<br/>
The wine dataset was a very clean dataset with no null values.I checked the dataset for inconstencies and set about exploring the data.Upon intial examination of the data, I found that all of the wines were rated between 3 and 9 though the scale is from 1 to 10.I started out by checking each variable individually, the summary command in R helped in giving me a sense of the spread of the data,I plotted histograms to understand the distribution of the data.The histograms showed fairly normal distributions with a few bimodal distributions and long tailed distributions,I realised that this was because of a few outliers in the data.I cleaned the data by eliminating the outliers, the datapoints above the 99th percentile and 0.01 percentile were removed.The histograms were more clear and the distributions were nearly normal distributons.The peaks of the histograms gave me information of the central tendencies of every variable visually.<br/>
I then looked at correlations between variables.The ggally package serves as an excellent package to generate scatterplot of correlations between all the variables.The scatterplot also showed correlation values.I noted the pairs with high correlation values and plotted them to get a better understanding of the trends,I was able to draw smoothing lines using geom_smooth which captured the trend clearly.<br/>
I was also interested in the effects of attributes on the quality.I plotted all the attribute quality pairs along with their boxplots to see the relationship.I compared the medians across all qualities with the over all median of the data related to each attribute.I could find distinct trends between attributes and quality.<br/>
Multivariate analysis was challenging as it is not easy to view trends with the original data so I reshaped data by cutting a few variables into different classes.The trends were clearly visible after reshaping,it also gave me more insight into wine preparation processes.<br/>
The other challenge I faced was the inavaialability of data corresponding to higher and lower qualities of wine.There might be more trends and patterns which I cannot discern with the present data set.<br/>
My analysis only dealt with highly correlated attributes,there could be trends and patterns hidden within uncorrelated attributes and also effects of certain attributes on quality could be much more than expected.This inference is supported by a few attributes getting surprisingly heavier weights on the linear model developed.<br/>
Overall the exercise helped me in develop insight about the data which I knew very little about simply by plotting it. 

